The error in Bayesian phylogenetic reconstruction when speciation is not instantaneous

### 2018-05-31 article.tex
After meeting @rsetienne

% @rsetienne:
% At the moment, there are 87 parameter combinations for one site and one clock model
% This allows for 40K / 87 = 460 replicates [*]
% Using two site and two clock models would result in approx 100 replicates
% Using four site and four clock models would result in approx 25 replicates
%
% I personally think the broadest data set will be most interesting,
% because it gives a rough idea, but -more importantly- shows the scientific
% community it can easily be done nowadays.
%
% [*] the number of replicates can only be determined by actually running the
% experiment, as there are multiple constraints that I cannot test for

@@

From an alignment, we run a Bayesian analysis and 
create a posterior distribution of trees and parameters
using the \verb;babette; (\cite{babette}) package
that sets the input parameters similar to BEAUti 2 and then runs BEAST2. 
For our site model, we assume either a Jukes-Cantor or GTR nucleotide substitution model.
The Jukes-Cantor model is the correct one, as it is used for simulating that alignment,
but the GTR model is the site model with the fewest assumptions and most flexibility.
\textbf{[NOTE: Why then not use all four site models?]}
For our clock model, we assume either a strict or relaxed log-normal \textbf{[NOTE: 
Why that one? Why then not use all clock models?]} clock model. The strict clock
model is the correct one, as it is used for simulating that alignment,
but the relaxed log-normal clock model is the model with fewer assumptions and
more flexibility.
We set the BD model as a tree prior, 
as gauging the effect of this incorrect assumption is the goal of this study. 
We assume an MRCA prior with a tight normal distribution
around the crown age, by choosing the crown age as mean, and a standard deviation 
of $0.5 \cdot 10^{-3}$ time units,
resulting in 95\% of the crown ages inferred have the same resolution (of $10^{-3}$ time 
units) as the alignment. 
We ran the MCMC chain to generate 1111 states,
of which we remove the first 10\% (also called the 'burn-in'). 
Of the remaining
1000 MCMC states, the effective sample size (ESS) of the posterior \textbf{[NOTE: I chose 
the ESS of the posterior over the ESS of the tree likelihood (and the 
others displayed in table 4). For both something can be said. Agree on this choice?]} must at least be 200
for a strong enough inference (\cite{beastbook}). An ESS can be increased by increasing
the number of samples or decreasing the autocorrelation between samples. 
If the ESS is less than 200, we decrease autocorrelation by doubling 
the MCMC sampling interval of that simulation, until the ESS exceeds 200.

>>>>

@@

From an alignment, we run a Bayesian analysis and 
create a posterior distribution of trees and parameters
using the \verb;babette; (\cite{babette}) package
that sets the input parameters similar to BEAUti 2 and then runs BEAST2. 
For our site model, we assume either a Jukes-Cantor or GTR nucleotide substitution model.
The Jukes-Cantor model is the correct one, as it is used for simulating that alignment,
where the GTR model is the site model that is picked as a default by most users.
For our clock model, we assume either a strict or relaxed log-normal 
clock model. 
Also here, the strict clock model is the correct one, as it is used for simulating that alignment,
but the relaxed log-normal clock model is the the one most commonly used.
We set the BD model as a tree prior, 
as gauging the effect of this incorrect assumption is the goal of this study. 
We assume an MRCA prior with a tight normal distribution
around the crown age, by choosing the crown age as mean, and a standard deviation 
of $0.5 \cdot 10^{-3}$ time units,
resulting in 95\% of the crown ages inferred have the same resolution (of $10^{-3}$ time 
units) as the alignment. 
We ran the MCMC chain to generate 1111 states,
of which we remove the first 10\% (also called the 'burn-in'). 
Of the remaining
1000 MCMC states, the effective sample size (ESS) of the posterior \textbf{[NOTE: I chose 
the ESS of the posterior over the ESS of the tree likelihood (and the 
others displayed in table 4). For both something can be said. Agree on this choice?]} must at least be 200
for a strong enough inference (\cite{beastbook}). An ESS can be increased by increasing
the number of samples or decreasing the autocorrelation between samples. 
If the ESS is less than 200, we decrease autocorrelation by doubling 
the MCMC sampling interval of that simulation, until the ESS exceeds 200.
